---
title: Project Airbnb
date: '2017-10-31T22:26:09-05:00'
description: Data wrangling with with real-world datasets
draft: no
image: 
keywords: ''
slug: Airbnb
categories:
- ''
- ''
---

```{r setup, include=FALSE}
options(knitr.table.format = "html") 
knitr::opts_chunk$set(warning = FALSE, message = FALSE, 
  comment = NA, dpi = 300)
```


```{r load-libraries, echo=FALSE}

library(tidyverse) # the usual stuff: dplyr, readr, and other goodies
library(lubridate) # to handle dates
library(GGally) # for correlation-scatter plot matrix
library(ggfortify) # to produce residual diagnostic plots
library(rsample) # to split dataframe in training- & testing sets
library(janitor) # clean_names()
library(broom) # use broom:augment() to get tidy table with regression output, residuals, etc
library(huxtable) # to get summary table of all models produced
library(kableExtra) # for formatting tables
library(moderndive) # for getting regression tables
library(skimr) # for skim
library(mosaic)
library(leaflet) # for interactive HTML maps
library(tidytext)
library(viridis)
library(vroom)
library(rsample)
library(car)
```


# Executive Summary

### The aim of the analysis is to build a prediction model that can estimate the average cost for 2 people to stay in an AirBnB in Amsterdam for 4 nights. As such, we analysed data about Airbnb listings in the city and identified predictors that are highly relevant to the price and fitted several models. Those predictors are selected first based on common sense and were analysed in terms of their price expanatory power. We then tested several models in order to find out which could predict the price better. The best model was chosen based on how well it helps to predict the prices given different inputs. The price is then estimated based on the best model. The best model includes variables such as property type, number_of_reviews, review_scores_rating, bathrooms, bedrooms, accommodates, host_is_superhost, neighbourhood_simplified, availability_30 and reviews_per_month. These are consistent with common sense metrics that would lead to increases or decreases in price.

```{r load_data, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}

listings <- vroom("http://data.insideairbnb.com/the-netherlands/north-holland/amsterdam/2021-09-07/data/listings.csv.gz") %>% 
       clean_names()

```

# Exploratory Data Analysis (EDA)

```{r, out.width='60%'}
#Storing the price as a number
listings_clean <- listings %>% 
  filter(name != "",
         host_location != "") %>% 
  mutate(price = parse_number(price),
         bathrooms = parse_number(gsub( "\\s.*", "", bathrooms_text))) %>% 
  filter(bathrooms != "")

```

> We noticed that for each variable there are a lot of outliers outside where the majority data concentrate and we would like to remove those uncommon observations later.

```{r}
# assumptions of the followings variables are made in accordance to common choices for tourists
listings_clean1 <- listings_clean %>% filter(bedrooms < 5) 
listings_clean3 <- listings_clean1 %>% filter (minimum_nights <= 4)
listings_clean4 <- listings_clean3 %>% filter (price < 1500) # Everything above that would be extraordinarily expensive
listings_clean5 <- listings_clean4 %>% filter (accommodates > 1)

# cleaning the data and choosing the relevant variables
listings_smaller <- listings_clean5 %>% 
  select(host_location,
         neighbourhood,
         neighbourhood_cleansed,
         neighbourhood_group_cleansed,
         property_type,
         room_type,
         accommodates,
         bathrooms,
         bathrooms_text,
         bedrooms,
         beds,
         price,
         review_scores_rating,
         minimum_nights,
         maximum_nights,
         property_type,
         latitude,
         longitude,
         host_is_superhost,
         instant_bookable,
         availability_30,
         reviews_per_month)
```
## Key Explanatory Variables
```{r}
# investigate the correlation among some key variables
listings_smaller %>% 
  select(price, bedrooms, beds, bathrooms, review_scores_rating) %>% 
  ggpairs(aes(alpha = 0.4))

```

```{r}

#creating a simplified version of `property_type`
listings_clean5 <- listings_clean5 %>%
  mutate(prop_type_simplified = case_when(
    property_type %in% c("Entire rental unit","Entire residential home", "Entire townhouse","Private room in rental unit") ~ property_type, 
    TRUE ~ "Other"
  ))
  
```

> We also noticed that the four most common property types in Airbnb Amsterdam are "Entire rental unit", "private room in rental unit", "entire condo" and "entire residential home". Collectively they account for the majority of the property types available in Amsterdam. The significantly high number of row houses, condos and houses with lofts relative to other major cities like Rotterdam (more metropolitan with numerous skyscapers) also justifies the relatively high number of condos and lofts.

> The most common values appearing for the variable "minimum nights" are 1,2 and 3 nights. Out of these 3 values, 2 nights is by far the most commonly occuring value. Given the fact that Amsterdam is amongst one of the top cities for a weekend visit by tourists residing in UK and western Europe, a minimum of 2 nights is seems logical. Similarly, 30 nights is the 9th highest with it being only 1 of 2 values to be greater than 10. The fact that any reservation >29 days is considered as a "long-term stay" and the host can earn significant benefits from the same could explain why a minimum of 30 nights is so common.  

# Mapping 

```{r, out.width = '60%'}

leaflet(data = filter(listings_clean5)) %>% 
  addProviderTiles("OpenStreetMap.Mapnik") %>% 
  addCircleMarkers(lng = ~longitude, 
                   lat = ~latitude, 
                   radius = 1, 
                   fillColor = "blue", 
                   fillOpacity = 0.4, 
                   popup = ~listing_url,
                   label = ~property_type)
```

> The concentration of the listings are around central amsterdam and south-east Amsterdam. This is an expected result given that central and south-east amsterdam tend to the key tourist spots like the "nine streets" and thus justifies the greater concentration of listings in this areas.  

# Regression Analysis


```{r, out.width = '50%'}

listings_clean5 <- listings_clean5 %>% 
  mutate(price_4_nights = price * 4) %>%  # we did not include accomodates because the given price is already including all people
  filter(review_scores_rating > 0) # we exclude 0's because new listing usually don't have reviews and are therefore not informative
  

ggplot(listings_clean5, aes(x = price_4_nights)) +
  geom_density()

ggplot(listings_clean5, aes(x = log(price_4_nights))) +
  geom_density()

```

> The variable (log(price_4_nights)) has been used as it effectively captures the compunding effect; it also allows for an earier and more meaningful interpretation of the regression results (in percentage terms). This is also showcased by the density plots; the variable price_4_nights is right skewed whereas the logged variable reflects a normal distribution bell curve.

## Model 3

```{r, out.width = '50%'}

model3 <- lm(log(price_4_nights) ~
               prop_type_simplified +
               number_of_reviews +
               review_scores_rating +
               # room_type + GVIF higher than 5
               bathrooms +
               bedrooms +
               beds +
               accommodates +
               host_is_superhost +
               instant_bookable,
             data = listings_clean5)

summary(model3)

vif(model3)


listings_clean5 <- listings_clean5 %>% 
  mutate(neighbourhood_simplified = case_when(neighbourhood_cleansed %in% c("Oostelijk Havengebied - Indische Buurt",
                                                                             "Oud-Oost",
                                                                             "Watergraafsmeer",
                                                                             "Noord-Oost",
                                                                             "IJburg - Zeeburgereiland",
                                                                             "IJburg - Zeeburgereiland",
                                                                             "Gaasperdam - Driemond",
                                                                             "Bijlmer-Oost")~"East",
                                              neighbourhood_cleansed %in% c("Centrum-Oost",
                                                                            "Centrum-West",
                                                                            "De Pijp - Rivierenbuurt",
                                                                            "Bijlmer-Centrum")~"Center",
                                              neighbourhood_cleansed %in% c("De Baarsjes - Oud-West",
                                                                            "Westerpark",
                                                                            "Slotervaart",
                                                                            "Bos en Lommer",
                                                                            "De Aker - Nieuw Sloten",
                                                                            "Osdorp",
                                                                            "Geuzenveld - Slotermeer")~"West",
                                              neighbourhood_cleansed %in% c("Zuid",
                                                                            "Buitenveldert - Zuidas")~"South",
                                              neighbourhood_cleansed %in% c("Noord-West",
                                                                            "Oud-Noord")~"North"))
                                                
                                              
unique(listings_clean$neighbourhood_simplified)
```

## Model 5
```{r, out.width = '50%'}
model5 <- lm(log(price_4_nights) ~
               prop_type_simplified +
               number_of_reviews +
               review_scores_rating +
               # room_type + GVIF higher than 5
               bathrooms +
               bedrooms +
               beds +
               accommodates +
               host_is_superhost +
               instant_bookable +
               neighbourhood_simplified +
               availability_30 +
               reviews_per_month,
             data = listings_clean5)

summary(model5)

vif(model5)
autoplot(model5)



```

## Model 6
> Removed variable beds due to high correlation with bedrooms and variable instant_bookable which is insignificant in model 5.

```{r, out.width = '50%'}

model6 <- lm(log(price_4_nights) ~
               prop_type_simplified +
               number_of_reviews +
               review_scores_rating +
               # room_type + GVIF higher than 5
               bathrooms +
               bedrooms +
               accommodates +
               host_is_superhost +
               neighbourhood_simplified +
               availability_30 +
               reviews_per_month,
             data = listings_clean5)

summary(model6)

```

## Diagnostics, collinearity, summary tables


```{r, out.width = '50%'}

huxreg(model3, model5, model6)

# Prediction using model6; model 6 has by far the highest R^2 and is less prone to collinearity than model5

listings_selection <- listings_clean5 %>% 
  filter(prop_type_simplified == "Private room in rental unit",
         number_of_reviews >= 10,
         review_scores_rating / 5 >= 0.9)

set.seed(1234)

train_test_split <- initial_split(listings_selection, prop = 0.75)
listings_train <- training(train_test_split)
listings_test <- testing(train_test_split)

rmse_train <- listings_train %>% 
  mutate(predictions = predict(model6, .)) %>% 
  summarise(sqrt(sum(predictions - log(price_4_nights))**2/n())) %>% 
  pull()

rmse_train # 0.05610197

rmse_test <- listings_test %>% 
  mutate(predictions = predict(model6, .)) %>% 
  summarise(sqrt(sum(predictions - log(price_4_nights))**2/n())) %>% 
  pull()

rmse_test # 0.2981508


listings_specifications <- tibble(prop_type_simplified = "Private room in rental unit",
               number_of_reviews = 10,
               review_scores_rating = 0.9 * 5,
               # room_type + GVIF higher than 5
               bathrooms = 2,
               bedrooms = 2,
               accommodates = 4,
               host_is_superhost = TRUE,
               neighbourhood_simplified = "Center",
               availability_30 = 4,
               reviews_per_month = 5)

exp(predict(model6, newdata = listings_specifications, interval = "confidence"))

# 95% CI for price for our specifications between $545.66 and $589.72 for our trip


```

>Regression Analysis and summary

From the Analysis above we can see that model6 is the best model to use out of the 6 models. Model6 explains us the highest variance of 45.8% in Prices while the rest are much below that with Model3 explaining 37.5% of the variance , Model5 45.7%, Model2 22.7%, Model1 17.1% and Model4 only 3.1%. Furthermore, Model6 incorporates the highest number of significant variables in model making it more robust and Explanatory than any of the other models. 
Hence, due to the aforementioned reasons we believe that Model6 is the best Model that should be used for further analysis.

> The number of bathrooms is significant at the 1% level and "accomodates" is significant at the 0.1% level. Similarly, the variable "bedroom" is also significant at the 5% level in model 5. This indicates that the independant variables do improve the regression model and thus allow for a more sound prediction of the expected price. The accomodation variable can be interpreted as - an increase in the individuals residing leads to a 14.9% rise in price. Similarly, an increase in the bedroom and bathroom leads to a 10.2% and 11.9% increase in the relevant price respectively. This is logical given that more 2 bedrooms corressponds to a larger flat/hotel and thus is likely to cost more.
Controlling for "host being a super-host" also suggests that a super-host tends to charge a  5.6% higher price than a normal host, however, this coefficient is statistically insignificant. Analogous to being a super host, the variable "instant book" is also statistically insiginicant at the 5% level and hence does not appear to have a large impact on the dependant variable (price_4_nights)
The 4 variables "neighbourhood simplified south, north, east and west" are streamlined to capture the effect of location on the dependant variable. As showcased in the table above, all 4 variables are statistically significant at the 0.1% level. The coefficients are all negative which suggests that, for example, staying in a listing located in the north leads to a 28.1% decline in expected price for 4 nights. This also appears to be logical as, people using airbnb in Amsterdam for 4 nights are likely to be tourists and thus will target a listing in central Amsterdam. This lack of demand for staying outside the central region indicates the negative effect of location on price.
Lastly, the variable "availability_30" is also statistically significant at the 0.1% level and thus suggests that the availability of the listing does have an impact on the listings price. Surpirisngly, the impact of the variable "listings_per_month" is insignificant and thus does not appear to have a commanding impact on the listings price. This could be due to several factors - tourists may be more concerned with the convenience and the location of the listing rather than the reviews, similarly, tourists could also potentially use Amsterdam as a 'gateway to western Europe' and thus the reviews are unlikely to have a significant impact on the listings price

